---
title: "Analysis of Divvy Trips 2019 Q1"
author: "Hanim Sofiah"
date: "2024-02-28"
output:
  pdf_document: default
  html_document: default
---

# Introduction

This document presents an analysis of Divvy Trips data for the first quarter of 2019. The analysis includes descriptive statistics, average ride lengths by user type and day of the week, and the number of rides by day of the week.

# Setup Environment

```{r setup}
library(dplyr)
library(lubridate)
library(readr)
library(ggplot2)
```

# Load Data

```{r load-data}
trips_data <- read_csv("Divvy_Trips_2019_Q1.csv")
```

# Descriptive Analysis

```{r descriptive-analysis}
summary(trips_data)
```

# Mean, Max and Mode

```{r calculate-stats}
mean_ride_length <- mean(trips_data$ride_length, na.rm = TRUE)
max_ride_length <- max(trips_data$ride_length, na.rm = TRUE)

getmode <- function(v) {
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v, uniqv)))]
}
mode_day_of_week <- getmode(trips_data$day_of_week)
```

# Average Ride Length by User Type and Day Of Week

```{r avg-ride-length}
avg_ride_length_by_type <- trips_data %>%
  group_by(usertype) %>%
  summarise(average_ride_length = mean(ride_length, na.rm = TRUE))

avg_ride_length_by_day_and_type <- trips_data %>%
  group_by(day_of_week, usertype) %>%
  summarise(average_ride_length = mean(ride_length, na.rm = TRUE))

rides_by_day <- trips_data %>%
  group_by(day_of_week) %>%
  summarise(number_of_rides = n())
```

# Visualization

## Average Ride Length by Type of User

```{r avg-ride-length-plot}
ggplot(avg_ride_length_by_type, aes(x = usertype, y = average_ride_length, fill = usertype)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Ride Length by User Type", x = "User Type", y = "Average Ride Length (minutes)") +
  theme_minimal()
```

## Average Ride Length by Day and Type

```{r avg-ride-length-by-day-plot}
ggplot(avg_ride_length_by_day_and_type, aes(x = day_of_week, y = average_ride_length, fill = usertype)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Average Ride Length by Day of Week and User Type", x = "Day of Week", y = "Average Ride Length (minutes)") +
  theme_minimal()
```

## Number of Rides by Day of the Week

```{r rides-by-day-plot}
ggplot(rides_by_day, aes(x = day_of_week, y = number_of_rides)) +
  geom_line(group=1) +
  labs(title = "Number of Rides by Day of the Week", x = "Day of Week", y = "Number of Rides") +
  theme_minimal()
```

